## Bionergy - Demand Forecast and Technology Evolution

##### [Oscar Jaimes](https://www.linkedin.com/in/oscj/) 

This set of graphs is designed to show how the demand for biofuels and technological advancements are linked in Canada's key bioenergy-producing regions, under the three CER future scenarios. These visuals are aimed at providing an estimate of the bioenergy sector's projected growth within each selected region. I'm closely following Alberta's progress as it diversifies its energy portfolio, adding to its traditional focus on oil and gas to include more renewable energy sources. This transition is not only important for environmental sustainability but also offers an opportunity for investors to be part of the expanding bioenergy market in Canada. Biofuel demand in Alberta is expected to approximately double by 2050 under the Canada Net-zero scenario from 2023 levels. This growth is exciting, and definitely worth following in the near future.

Sources: 

- [Canada's Energy Future 2023: Energy Supply and Demand Projections, End Use Demand](https://open.canada.ca/data/en/dataset/7643c948-d661-4d90-ab91-e9ac732fc737/resource/9003d40e-087b-4af0-b2c8-2217fd697a28)
- [Canada's Energy Future 2023: Energy Supply and Demand Projections, Electricity Generation Technology](https://open.canada.ca/data/en/dataset/7643c948-d661-4d90-ab91-e9ac732fc737/resource/55ebd8ad-20b8-4730-b6d0-0b1c6a480aea)


```{r packages-data,echo=F,output=F, warning=FALSE,include=F}
library(kableExtra)
library(readxl)
library(janitor)
library(tidyverse)
library(lubridate)
library(scales)
library(viridis)
library(gridExtra)
library(wesanderson)
knitr::opts_chunk$set(echo=FALSE)
```


```{r chunk dataload,warning=FALSE,echo=FALSE,include=FALSE}

end_use <- read_csv(
  'https://www.cer-rec.gc.ca/open/energy/energyfutures2023/end-use-demand-2023.csv',
  show_col_types = FALSE
  ) %>% clean_names()
  
technology <- read_csv(
  'https://www.cer-rec.gc.ca/open/energy/energyfutures2023/electricity-generation-technology-2023.csv',
  show_col_types = FALSE
  ) %>% clean_names()

technology$variable[technology$variable == "Bioenergy with CCUS"] <- "Bioenergy with CCUS (Carbon Capture, Utilisation, and Storage)"
```

```{r chunk title,warning=FALSE,echo=FALSE,include=FALSE}
generate_plots_for_region <- function(r) {
  end_use_demand <-end_use %>%
    filter(region == r) %>%
    filter(variable == "Biofuels & Emerging Energy") %>%
    filter(sector != "Total End-Use") %>%
    filter(year <= 2023)
  
  end_use_demand_proj <- end_use %>%
    filter(region == r) %>%
    filter(variable == "Biofuels & Emerging Energy") %>%
    filter(sector != "Total End-Use") %>%
    filter(year >= 2023)
  
  electricity_technology <- technology %>%
    filter(region == r) %>%
    filter(variable %in% c("Bioenergy", "Bioenergy with CCUS (Carbon Capture, Utilisation, and Storage)")) %>%
    filter(year <= 2023)
  
  electricity_technology_proj <- technology %>%
    filter(region == r) %>%
    filter(variable %in% c("Bioenergy", "Bioenergy with CCUS (Carbon Capture, Utilisation, and Storage)")) %>%
    filter(year >= 2023)
  
  end_use_p <- end_use_demand %>%
    ggplot()+
    scale_fill_manual(values = c( "#909191", "#446455", "#fbd364"))+
    theme_minimal()+ 
    geom_area(
      aes(year,value, group=sector, fill=sector), 
      position = "stack", 
      colour="black",
      linewidth=0.5
    )+
    geom_area(
      data=end_use_demand_proj,
      aes(year,value, group=sector, fill=sector), 
      position = "stack", 
      colour="black",
      linewidth=0.5,
      alpha=0.5
    )+
    theme(
      text = element_text(size=12, face="bold"),
      legend.position = "bottom",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.subtitle = element_text(hjust = 0.0, size=10),
      strip.background = element_rect(fill = "white", color = "black", size = 1.5),
      strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10),
      plot.caption = element_text(hjust = 0, vjust = 1) 
    )+
    geom_vline(xintercept=2023, linetype="dotted")+
    guides(fill=guide_legend(nrow=1, title=NULL))+
    facet_wrap(~scenario)+
    labs(
      title="End-Demand For Biofuels, by Sector",
      y="Petajoules",x="",
      subtitle=("CER End-Demand Data (2005-2023) and Projections (2023-2050)")
    )
  
  
  technology_p <- electricity_technology %>%
    ggplot()+
    scale_fill_manual(values = wes_palette("IsleofDogs1"))+
    theme_minimal()+ 
    geom_area(
      aes(year,value*0.0036, group=variable, fill=variable), 
      position = "stack", 
      colour="black",
      linewidth=0.5
    )+
    geom_area(
      data=electricity_technology_proj,
      aes(year,value*0.0036, group=variable, fill=variable), 
      position = "stack", 
      colour="black",
      linewidth=0.5,
      alpha=0.65
    )+
    theme(
      text = element_text(size=12, face="bold"),
      legend.position = "bottom",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      plot.subtitle = element_text(hjust = 0.0, size=10),
      strip.background = element_rect(fill = "white", color = "black", size = 1.5),
      strip.text.x = element_text(size = 10),
      strip.text.y = element_text(size = 10),
      plot.caption = element_text(hjust = 0, vjust = 1, face="italic") 
    )+
    geom_vline(xintercept=2023, linetype="dotted")+
    guides(fill=guide_legend(nrow=1, title=NULL))+
    facet_wrap(~scenario)+
    labs(
      title="Increase in Bioenergy Generation From Technology Evolution",
      y="Petajoules",x="",
      subtitle=("CER Generation Data (2020-2023) and Projections (2023-2050)"),
      caption="Source: CER Canada's Energy Futures 2023, Graph by Oscar Jaimes"
    )


  return(list(end_use_p, technology_p))

}

province_names = c("Alberta", "British Columbia", "Ontario", "Quebec", "Canada")
plotly_plots = lapply(province_names, generate_plots_for_region)
```



##  {.tabset}

```{r biofules_and_energy,fig.width=12,fig.height=4.5,dpi=200, results='asis'}
for (i in 1:length(plotly_plots)) {
  cat("### ",province_names[i],"\n")
  print(plotly_plots[[i]][[1]])
  cat('\n\n')
  print(plotly_plots[[i]][[2]])
  cat('\n\n')
}
```

## 

My name is Oscar, and I currently in my final semester at the University of Alberta, pursuing a major in Computing Science. My academic journey has been enhanced with unique work placements in software/data engineering. In 2022, I completed an 8-month term at BlackBerry in Ottawa, contributing to their [Radar platform](https://www.blackberry.com/us/en/products/asset-tracking). My most significant experience, however, has been with [Livestock Water Recycling (LWR)](https://www.livestockwaterrecycling.com), a leading agTech company based in Calgary. Since my second year (2019), I have progressed to a full-time software engineer at LWR, leading the development of data pipelines and software solutions for manure and wastewater treatment systems. LWR's expertise in manure treatment technology notably contributes to the Renewable Natural Gas (RNG) sector, supplying valuable feedstock for anaerobic digesters and facilitating a more sustainable RNG lifecycle. Feel free to checkout my [GitHub](https://github.com/oscj) and [LinkedIn](https://www.linkedin.com/in/oscj/) for a deeper look into my projects and experience. 
