---
title: "Data Assignment 3"
author: "Oscar Jaimes"
date: "2023-11-26"
output: html_document
format: 
  html:
    code-fold: true
    page-layout: "full"
toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=F,warning=F)
options(scipen = 999)

library(kableExtra)
library(readxl)
library(janitor)
library(tidyverse)
library(lubridate)
library(scales)
library(viridis)
library(cansim)
library(cowplot)
library(gridExtra)
library(wesanderson)
```


## Electricity Generation by Source, Canada

### Download Data
```{r}
if(!file.exists("prov_elec.xlsx")){
   download.file(
     "https://econ366.aleach.ca/resources/assignment_3/prov_elec.xlsx",
     destfile = "prov_elec.xlsx",
     mode="wb"
    )
}
can_elec <- read_excel("prov_elec.xlsx", sheet="Table A13-1", range="A1:T13")
```

### Parse Data
```{r}
# wide to long
can_elec <- pivot_longer(can_elec, cols = -1, names_to = "Year", values_to ="Generation")
can_elec <- setNames(can_elec, c("Source", "Year", "Generation"))

# same order as assignment description
desired_order <- c("Coal", "Natural Gas", "Other Fuels", "Nuclear", "Hydro", "Other Renewables", "Other Generation")
can_elec$Source <- factor(can_elec$Source, levels = desired_order)
```


### Graph
```{r, fig.width=12, fig.height=7}
# these palettes are too small to represent each source by themselves, so combine them into a single palette
combined_palette <- c(wes_palette("Chevalier1"), wes_palette("Moonrise2"))
can_elec %>%
  ggplot()+
  scale_fill_manual(values = combined_palette)+
  # divide Generation by 1000 to get TWh instead of GWh
  geom_col(aes(Year,Generation/1000,fill=Source),color="black",linewidth=.25)+
  guides(fill=guide_legend(nrow = 1))+
  theme_minimal()+ 
  theme(
    text = element_text(size=16),
    axis.text.x=element_text(angle=45,vjust=1,hjust=1),
    legend.position = "bottom"
  )+
  guides(fill=guide_legend(nrow=1, title=NULL))+
  scale_y_continuous(
    breaks = seq(0, 600, by = 100),
  )+
  labs(
    y="Electricity Generation (TWh)",
    x="",
    title="National Electricity Generation by Source",
    caption="Source: Canada's National GHG Emissions Inventory, 2022, graph by Oscar Jaimes"
  )
```


## Electricity Generation by Source, Alberta

### Read Data
```{r}
ab_elec <- read_excel("prov_elec.xlsx", sheet="Table A13-10", range="A1:T8")
```

### Parse Data
```{r}
# wide to long
ab_elec <- pivot_longer(ab_elec, cols = -1, names_to = "Year", values_to ="Generation")
ab_elec <- setNames(ab_elec, c("Source", "Year", "Generation"))

# same order as assignment description
desired_order <- c("Coal", "Natural Gas", "Other Fuels", "Nuclear", "Hydro", "Other Renewables", "Other Generation")
ab_elec$Source <- factor(ab_elec$Source, levels = desired_order)
```


### Graph
```{r, fig.width=12, fig.height=7}
ab_elec %>%
  ggplot()+
  scale_fill_manual(values = combined_palette)+
  # divide Generation by 1000 to get TWh instead of GWh
  geom_col(aes(Year,Generation/1000,fill=Source),color="black",linewidth=.25)+
  guides(fill=guide_legend(nrow = 1))+
  theme_minimal()+ 
  theme(
    text = element_text(size=16),
    axis.text.x=element_text(angle=45,vjust=1,hjust=1),
    legend.position = "bottom"
  )+
  guides(fill=guide_legend(nrow=1, title=NULL))+
  labs(
    y="Electricity Generation (TWh)",
    x="",
    title="Electricity Generation by Source, Alberta",
    caption="Source: Canada's National GHG Emissions Inventory, 2022, graph by Oscar Jaimes"
  )
```

## Emissions from Electricity Generation, Canada

### Download Data
```{r}
if(!file.exists("emission_projections.csv")){
   download.file(
     "https://econ366.aleach.ca/resources/assignment_3/assignment_3_projections.csv",
     destfile = "emission_projections.csv",
     mode="wb"
    )
}

proj_data <- read_csv("emission_projections.csv")
```

### Parse Data
```{r}
nir_data <- proj_data %>%
  filter(scenario == "NIR 2022")%>%
  filter(!(prov == "Canada"))%>%
  filter(year <= 2020)%>%
  filter(emissions > 0) %>%
  mutate(prov=fct_relevel(prov,"Canada","BC","AB" ,"SK","MB", "ON","QC","ATL", "Northwest Territories"))
```

### Graph 1 - Canadian Electricity Sector GHG Emissions by Province
```{r, fig.width=12, fig.height=7}
custom_palette <- c("#00204e", "#233f6c", "#575c6d", "#7c7b79","#a69e75", "#d3c164", "#ffea46", "#ffea46")

nir_data %>%
 ggplot()+
 scale_fill_manual(values = custom_palette)+
  geom_area(
    aes(year,emissions,group=prov,fill=prov),
    position = "stack",
    colour="black",
    linewidth=0.2
  )+
  theme_minimal()+
  theme(
    text = element_text(size=14, face="bold"),
    plot.caption = element_text(hjust = 0, vjust = 1, face="italic"),
  )+
  scale_y_continuous(
    breaks = seq(0, 125, by = 25),
  )+
  scale_x_continuous(
    breaks=pretty_breaks(n=3),
    expand=c(0,0),
  )+
  guides(fill=guide_legend(ncol=1, title=NULL))+
  labs(
    y="Annual Emissions (MtCO2e)",
    x="",
    title="Canadian Electricity Sector GHG Emissions by Province",
    caption="Source: Environment and Climate Change Canada, graph by Oscar Jaimes"
  )
```

### Graph 2 - Canadian Electricity Sector GHG Emissions and Projections by Province

```{r}
# Filter out initial data for 2022 reference case
proj_data <- proj_data %>%
  filter(scenario == "2022 Reference Case")%>%
  filter(emissions > 0)%>%
  filter(!(prov == "Canada"))%>%
  mutate(prov=fct_relevel(prov,"Canada","BC","AB" ,"SK","MB", "ON","QC","ATL", "Northwest Territories"))

# get the 2020 data from NIR 2022 and join it to the projection dataframe so there
# is no gap in the graph between the inventory and projection
nir_2020_data <- nir_data %>% filter(year == 2020)
proj_data <- bind_rows(proj_data, nir_2020_data)
```

```{r, fig.width=12, fig.height=7}
nir_data %>%
  ggplot()+ #make a graph
  scale_fill_manual(values = custom_palette)+
  geom_area(
    aes(year,emissions,group=prov,fill=prov),
    position = "stack",
    colour="black",
    size=0.1,
    linewidth=0.2,
  )+
  geom_area(
    data=proj_data,
    aes(year,emissions,group=prov,fill=prov),
    position = "stack",
    size=0.1,
    colour="black",
    linewidth=0.2,
    alpha=0.45
  )+
  scale_x_continuous(
    breaks=pretty_breaks(n=6),
    expand=c(0,0),
  )+
  theme_minimal()+
  theme(
    text = element_text(size=14, face="bold"),
    plot.caption = element_text(hjust = 0, vjust = 1, face="italic"),
    plot.subtitle = element_text(size=12)
  )+
  geom_vline(xintercept=2020, linetype="dotted")+
  scale_y_continuous(
    breaks = seq(0, 120, by = 20),
  )+
  guides(fill=guide_legend(ncol=1, title=NULL))+
  labs(
    y="Annual Emissions (MtCO2e)",
    x="",
    title="Canadian Electricity Sector GHG Emissions",
    subtitle="2022 National Inventory (2005-2020) levels and 2022 Reference Case projections (2020-2035, lighter fill)",
    caption="Source: Environment and Climate Change Canada, graph by Oscar Jaimes"
  )
```


## Electricity Microgeneration

### Download Data
```{r}
if(!file.exists("micro_data.xlsx")){
   download.file(
     "https://www.aeso.ca/assets/Uploads/market-and-system-reporting/Small-DERS-Microgen-Monthly-by-Type.xlsx",
     destfile = "micro_data.xlsx",
     mode="wb"
    )
}

microgen <- read_excel("micro_data.xlsx", sheet="Microgen")
```


### Parse & Filter
```{r}
microgen <- microgen %>% clean_names ()

microgen$microgen_fuel_type <- as_factor(microgen$microgen_fuel_type)
microgen$microgen_fuel_type <- fct_other(microgen$microgen_fuel_type, keep = "Solar", other_level = "Other")
microgen$month <- as.Date(microgen$month, format = "%Y-%d-%m")

microgen <- microgen %>%
  group_by(month, microgen_fuel_type) %>%
  summarize(
    total_capacity_mw=sum(microgen_capacity_mw),
    total_count = sum(microgen_count)
  )
```

### Graph 1 - Total Capacity by Fuel Type
```{r, fig.width=12, fig.height=7}
palette_colors <- rev(wes_palette("Chevalier1")[1:2])
capacity <- microgen %>%
  ggplot()+
  scale_fill_manual(values =palette_colors)+
  geom_area(
    aes(month, total_capacity_mw, group=microgen_fuel_type, fill=microgen_fuel_type),
    position = "stack",
    colour="black",
    linewidth=0.2
  )+
  theme_minimal()+
  scale_x_date(
    breaks=pretty_breaks(n=8),
    expand=c(0,0),
  )+
  scale_y_continuous(
    expand=c(0,0)
  )+
  theme(
    legend.position = "bottom",
    text = element_text(size=14, face="bold"),
    plot.caption = element_text(hjust = 0, vjust = 1, face="italic"),
    plot.subtitle = element_text(size=12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )+
  guides(fill=guide_legend(nrow=1, title=NULL))+
  labs(
    y="Microgen Capacity (MW)",
    x="",
    title="AB Electricity Microgeneration Capacity (MW)",
  )

capacity
```


### Graph 2 - Total Sites by Fuel Type
```{r, fig.width=12, fig.height=7}
sites <- microgen %>%
  ggplot()+
  scale_fill_manual(values = palette_colors)+
  geom_area(
    aes(month, total_count, group=microgen_fuel_type, fill=microgen_fuel_type),
    position = "stack",
    colour="black",
    linewidth=0.2
  )+
  theme_minimal()+
  scale_x_date(
    breaks=pretty_breaks(n=8),
    expand=c(0,0),
  )+
  scale_y_continuous(
    expand=c(0,0),
    breaks = seq(0, 15000, by = 5000),
  )+
  theme(
    legend.position = "bottom",
    text = element_text(size=14, face="bold"),
    plot.caption = element_text(hjust = 0, vjust = 1, face="italic"),
    plot.subtitle = element_text(size=12),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
  )+
  guides(fill=guide_legend(nrow=1, title=NULL))+
  labs(
    y="Microgen Sites (#)",
    x="",
    title="AB Electricity Microgeneration Sites (#)",
  )

sites
```


### Graph 3 - Combined Microgen Capacity and Sites
```{r, fig.width=12, fig.height=7}
cowplot::plot_grid(
  capacity,
  sites
)
```




